# Question 1

We used the data corresponding to all departures from the Austin/Bergstrom Airport in 2008 to create a set 12 of heatmaps, one for each month of the year that depict the average departure delays(in hours) by aircraft carrier and day of the week.

These charts can help us to get a sense of which aircraft carriers are most delayed on each day of the week for every month. We can also observe how delay patterns changed over the year.

```{r setup, include=FALSE}
# loading all libraries
library(plyr)
library(dplyr)
library(ggplot2)
library(reshape)
library(gplots)
```

```{r warning=FALSE}
ABIA=read.csv('ABIA.csv')
ABIA_dep=ABIA[ABIA$Origin=='AUS',]
ABIA_dep_df=tbl_df(ABIA_dep)
ABIA_dep$DayOfWeek=as.factor(ABIA_dep$DayOfWeek)
ABIA_dep$DayOfWeek =revalue(ABIA_dep$DayOfWeek ,c("1"="Mon","2"="Tue","3"="Wed","4"="Thurs","5"="Fri","6"="Sat","7"="Sun"))
ABIA_dep$Month=as.factor(ABIA_dep$Month)
ABIA_dep$Month =revalue(ABIA_dep$Month ,c("1"="January","2"="February","3"="March","4"="April","5"="May","6"="June","7"="July",
    "8"="August","9"="September","10"="October","11"="November","12"="December"))
```

```{r warning=FALSE,message=FALSE}
color=colorRampPalette(c('white','red'))
for (month in levels(ABIA_dep$Month)){
                    ABIA_dep_month=ABIA_dep[ABIA_dep$Month==month,]
                    ABIA_dep_df=tbl_df(ABIA_dep_month)
                    carrier=group_by(ABIA_dep_df,UniqueCarrier, DayOfWeek)
                    carrier_delay = summarize(carrier, delay = mean(DepDelay, na.rm = TRUE))
                    carrier_de=cast(carrier_delay, UniqueCarrier~DayOfWeek)
                    heatmap.2(as.matrix(carrier_de),Rowv=NA, Colv=NA,na.rm=TRUE,main=month,col=color,trace='none')
                    cat("\n\n\n\n\n")
                    }
```

White cells represent no flights relevant to that particular cell.

